### 2024-06-25 basic-statistics.qmd
20240624 - rearrange for consistency

```{r}
library("petersenlab")
library("pwr")
library("WebPower")
library("grid")
library("tidyverse")
```

@@

@@

$$
\eta_p^2 = = \frac{SS_{\text{effect}}}{SS_{\text{effect}} + SS_{\text{error}}}
$$ {#eq-partialEtaSquared}

@@

```{r}
power <- .8
effectSize_d <- .5
effectSize_r <- .24
effectSize_f <- .25
effectSize_fSquared <- .06
```

@@

```{r}
pwr::pwr.t.test(
  d = effectSize_d,
  power = power,
  sig.level = alpha,
  type = "one.sample",
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.t.test(
  n = N,
  power = power,
  sig.level = alpha,
  type = "one.sample",
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.t.test(
  d = effectSize_d,
  power = power,
  sig.level = alpha,
  type = "two.sample",
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.t.test(
  n = N,
  power = power,
  sig.level = alpha,
  type = "two.sample",
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.t2n.test(
  n1 = nGroup1,
  d = effectSize_d,
  power = power,
  sig.level = alpha,
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.t2n.test(
  n1 = nGroup1,
  n2 = nGroup2,
  power = power,
  sig.level = alpha,
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.t.test(
  d = effectSize_d,
  power = power,
  sig.level = alpha,
  type = "paired",
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.t.test(
  n = N,
  power = power,
  sig.level = alpha,
  type = "paired",
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.anova.test(
  f = effectSize_f,
  power = power,
  sig.level = alpha,
  k = numGroups)
```

@@

```{r}
pwr::pwr.anova.test(
  n = N,
  power = power,
  sig.level = alpha,
  k = numGroups)
```

@@

```{r}
wp.rmanova(
  n = N,
  ng = numGroups,
  nm = numMeasurements,
  f = effectSize_f,
  alpha = alpha,
  type = 0)

wp.rmanova(
  n = N,
  ng = numGroups,
  nm = numMeasurements,
  f = effectSize_f,
  alpha = alpha,
  type = 1)

wp.rmanova(
  n = N,
  ng = numGroups,
  nm = numMeasurements,
  f = effectSize_f,
  alpha = alpha,
  type = 2)
```

@@

```{r}
wp.rmanova(
  power = power,
  ng = numGroups,
  nm = numMeasurements,
  f = effectSize_f,
  alpha = alpha,
  type = 0)

wp.rmanova(
  power = power,
  ng = numGroups,
  nm = numMeasurements,
  f = effectSize_f,
  alpha = alpha,
  type = 1)

wp.rmanova(
  power = power,
  ng = numGroups,
  nm = numMeasurements,
  f = effectSize_f,
  alpha = alpha,
  type = 2)
```

@@

```{r}
wp.rmanova(
  n = N,
  ng = numGroups,
  nm = numMeasurements,
  power = power,
  alpha = alpha,
  type = 0)

wp.rmanova(
  n = N,
  ng = numGroups,
  nm = numMeasurements,
  power = power,
  alpha = alpha,
  type = 1)

wp.rmanova(
  n = N,
  ng = numGroups,
  nm = numMeasurements,
  power = power,
  alpha = alpha,
  type = 2)
```

@@

```{r}
pwr::pwr.r.test(
  r = effectSize_r,
  power = power,
  sig.level = alpha,
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.r.test(
  n = N,
  power = power,
  sig.level = alpha,
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.f2.test(
  f2 = effectSize_fSquared,
  sig.level = alpha,
  u = numPredictors,
  v = N - numPredictors - 1)
```

Solving for sample size per group needed (given effect size, power, and alpha)—$v = N - \text{numberOfPredictors} - 1$; thus, $N = v + \text{numberOfPredictors} + 1$:

```{r}
multipleRegressionSampleSizeModel <- pwr::pwr.f2.test(
  f2 = effectSize_fSquared,
  power = power,
  sig.level = alpha,
  u = numPredictors)

@@

vNeeded <- multipleRegressionSampleSizeModel$v
sampleSizeNeeded <- vNeeded + numPredictors + 1
sampleSizeNeeded
```

@@

@@

>>>>

```{r}
library("petersenlab")
library("pwr")
library("pwrss")
library("WebPower")
library("grid")
library("tidyverse")
```

@@

To calculate the effect size of a particular predictor, you can calculate $\Delta f^2$ as in @eq-deltaCohensFsquared:

$$
\begin{aligned}
  \Delta f^2 &= \frac{R^2_{\text{model}} - R^2_{\text{reduced}}}{1 - R^2_{\text{model}}} \\
             &= \frac{\eta^2_{\text{model}} - \eta^2_{\text{reduced}}}{1 - \eta^2_{\text{model}}}
\end{aligned}
$$ {#eq-deltaCohensFsquared}

where $R^2_{\text{model}$ is the $R^2$ of the model with the [predictor variable](#sec-correlationalStudy) of interest and R^2_{\text{reduced}} is the $R^2$ of the model without the [predictor variable](#sec-correlationalStudy) of interest.

@@

$$
\eta_p^2 = \frac{SS_{\text{effect}}}{SS_{\text{effect}} + SS_{\text{error}}}
$$ {#eq-partialEtaSquared}

@@

```{r}
power <- .8
effectSize_d <- .5
effectSize_r <- .24
effectSize_beta <- .24
effectSize_f <- .25
effectSize_fSquared <- .06
effectSize_omega <- .3
```

@@

```{r}
pwr::pwr.t.test(
  power = power,
  d = effectSize_d,
  sig.level = alpha,
  type = "one.sample",
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.t.test(
  power = power,
  n = N,
  sig.level = alpha,
  type = "one.sample",
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.t.test(
  power = power,
  d = effectSize_d,
  sig.level = alpha,
  type = "two.sample",
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.t.test(
  power = power,
  n = N,
  sig.level = alpha,
  type = "two.sample",
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.t2n.test(
  power = power,
  n1 = nGroup1,
  d = effectSize_d,
  sig.level = alpha,
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.t2n.test(
  power = power,
  n1 = nGroup1,
  n2 = nGroup2,
  sig.level = alpha,
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.t.test(
  power = power,
  d = effectSize_d,
  sig.level = alpha,
  type = "paired",
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.t.test(
  power = power,
  n = N,
  sig.level = alpha,
  type = "paired",
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.anova.test(
  power = power,
  f = effectSize_f,
  sig.level = alpha,
  k = numGroups)
```

@@

```{r}
pwr::pwr.anova.test(
  power = power,
  n = N,
  sig.level = alpha,
  k = numGroups)
```

@@

```{r}
WebPower::wp.rmanova(
  n = N,
  ng = numGroups,
  nm = numMeasurements,
  f = effectSize_f,
  alpha = alpha,
  type = 0)

WebPower::wp.rmanova(
  n = N,
  ng = numGroups,
  nm = numMeasurements,
  f = effectSize_f,
  alpha = alpha,
  type = 1)

WebPower::wp.rmanova(
  n = N,
  ng = numGroups,
  nm = numMeasurements,
  f = effectSize_f,
  alpha = alpha,
  type = 2)
```

@@

```{r}
WebPower::wp.rmanova(
  power = power,
  ng = numGroups,
  nm = numMeasurements,
  f = effectSize_f,
  alpha = alpha,
  type = 0)

WebPower::wp.rmanova(
  power = power,
  ng = numGroups,
  nm = numMeasurements,
  f = effectSize_f,
  alpha = alpha,
  type = 1)

WebPower::wp.rmanova(
  power = power,
  ng = numGroups,
  nm = numMeasurements,
  f = effectSize_f,
  alpha = alpha,
  type = 2)
```

@@

```{r}
WebPower::wp.rmanova(
  power = power,
  n = N,
  ng = numGroups,
  nm = numMeasurements,
  alpha = alpha,
  type = 0)

WebPower::wp.rmanova(
  power = power,
  n = N,
  ng = numGroups,
  nm = numMeasurements,
  alpha = alpha,
  type = 1)

WebPower::wp.rmanova(
  power = power,
  n = N,
  ng = numGroups,
  nm = numMeasurements,
  alpha = alpha,
  type = 2)
```

@@

```{r}
pwr::pwr.r.test(
  power = power,
  r = effectSize_r,
  sig.level = alpha,
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.r.test(
  power = power,
  n = N,
  sig.level = alpha,
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.f2.test(
  f2 = effectSize_fSquared,
  sig.level = alpha,
  u = numPredictors,
  v = N - numPredictors - 1)

pwrss::pwrss.t.reg(
  n = N,
  beta1 = effectSize_beta,
  k = numPredictors,
  alpha = alpha,
  alternative = "not equal")
```

Solving for sample size needed (given effect size, power, and alpha)—$v = N - \text{numberOfPredictors} - 1$; thus, $N = v + \text{numberOfPredictors} + 1$:

```{r}
multipleRegressionSampleSizeModel <- pwr::pwr.f2.test(
  power = power,
  f2 = effectSize_fSquared,
  sig.level = alpha,
  u = numPredictors)

@@

vNeeded <- multipleRegressionSampleSizeModel$v
sampleSizeNeeded <- vNeeded + numPredictors + 1
sampleSizeNeeded

pwrss::pwrss.t.reg(
  power = power,
  beta1 = effectSize_beta,
  k = numPredictors,
  alpha = alpha,
  alternative = "not equal")
```

@@

```{r}
pwr::pwr.chisq.test(
  N = N,
  w = effectSize_omega,
  df = numCategories - 1,
  sig.level = alpha)
```

```{r}
pwr::pwr.chisq.test(
  power = power,
  w = effectSize_omega,
  df = numCategories - 1,
  sig.level = alpha)
```

```{r}
pwr::pwr.chisq.test(
  power = power,
  N = N,
  df = numCategories - 1,
  sig.level = alpha)
```

@@

```{r}
pwr::pwr.chisq.test(
  N = N,
  w = effectSize_omega,
  df = (numRows - 1)*(numColumns - 1),
  sig.level = alpha)
```

```{r}
pwr::pwr.chisq.test(
  power = power,
  w = effectSize_omega,
  df = (numRows - 1)*(numColumns - 1),
  sig.level = alpha)
```

```{r}
pwr::pwr.chisq.test(
  power = power,
  N = N,
  df = (numRows - 1)*(numColumns - 1),
  sig.level = alpha)
```
